Sequential algorithms to split and merge ultra-high resolution 3D images

### 2017-08-02 paper.tex
Inserted @ValHayot's results on SSD. Updates in text.

\begin{abstract}
  Splitting and merging data is a requirement for most parallel or
  distributed processing operations. Naive algorithms, however,
  perform very poorly to split or merge 3D blocks from ultra-high
  resolution images, due to important seek times. We propose and
  analyze two algorithms based on in-memory reconstruction of image
  parts. Cluster reads access image chunks only once while Multiple
  reads minimize the accesses to the image. Evaluation on a
  3000x3000x3000 brain image shows that our algorithms perform
  similarly to the optimal configuration (slice merging) provided that
  enough memory is available, which solves our initial problem.  We
  conclude that splitting and merging large 3D images can be done
  efficiently based on simple data formats.
\end{abstract}

@@

\todo{Make more simulations to support that claim.}

@@

\begin{figure}
  \includegraphics[width=0.45\columnwidth]{figures/model-big-brain-comparison.pdf}
  \includegraphics[width=0.45\columnwidth]{figures/model-big-brain-comparison-moreblocks.pdf}
  \caption{Number of seeks for Cluster reads vs Multiple reads when
    Cluster reads merge complete block rows (case 2), for D=3000 and
    b=2. Left: n=125; Right: n=64,000.}
  % Gnuplot file: scripts/model/model.gnplt
  \label{fig:model-comparison}
\end{figure}

@@

\subsection{Infrastructure}

@@

Figure~\ref{fig:slices-vs-blocks} shows the merging time for Slices vs
Blocks, on SSD and HDD. Merging with Blocks takes about twice the time
of Slices, which motivates the need for improved algorithms to merge
blocks. As expected, seek time is mainly responsible for this
difference, although differences in read times are also observed,
perhaps due to a caching effect at the disk or OS level. This behavior is
reproduced in SSD and HDD, which shows that SSDs might not be the only
solution to reduce seek time.  \todo{On figure: check that breakdowns
  add up to total time. Harmonize time formats. Harmonize colors (all
  slices in purple, all blocks in green). If time permits, try
  histogram for breakdown, on same figure than total time.}
\begin{figure}[h]
  \centering
  \includegraphics[width=0.45\columnwidth]{figures/blocks-slices/totalmergetimeSSD1.pdf}
  \hfill
  \includegraphics[width=0.45\columnwidth]{figures/blocks-slices/totalmergetimeHDD1.pdf}\\
  \includegraphics[width=0.45\columnwidth]{figures/blocks-slices/hddBreakDown1.pdf}
  \hfill
  \includegraphics[width=0.45\columnwidth]{figures/blocks-slices/ssdBreakDown1.pdf}
  \caption{Merging time for Slices and Blocks. Left column: SSD. Right column: HDD. Top row: overall merging time. Bottom row: breakdown by read, write and seek time.}
  % Gnuplot file: scripts/blocks-slices/createboxplot.gnplt
\label{fig:slices-vs-blocks}
\end{figure}

@@

@@

@@

\begin{figure}[h]
  \centering
  \includegraphics[width=0.45\columnwidth]{figures/benchmark-mreads/mreads-breakdown-ssd.pdf}
  \hfill
    \includegraphics[width=0.45\columnwidth]{figures/benchmark-mreads/mreads-breakdown-hdd.pdf}
  \caption{Comparison between Multiple reads, blocks and slices. Left: SSD. Right: HDD.}
  % Python file: scripts/mreads/generate_avg_var_data.py
  % Gnuplot file: scripts/mreads/compare_breakdown_errbar.gnuplot
\label{fig:multiple-reads}
\end{figure}

\begin{figure}[h]
  \centering
  \includegraphics[width=0.70\columnwidth]{figures/benchmark-mreads/mreads-number-of-seeks.pdf}
  \hfill
  \caption{Number of seeks in Multiple reads}
  % Gnuplot file: scripts/mreads/number_of_seeks.gnuplot
\label{fig:multiple-reads-number-of-seek}
\end{figure}

>>>>

\begin{abstract}
  Splitting and merging data is a requirement for most parallel or
  distributed processing operations. Naive algorithms, however,
  perform very poorly to split or merge 3D blocks from ultra-high
  resolution images, due to important seek times. We propose and
  analyze two algorithms based on in-memory reconstruction of image
  parts. Cluster reads access image chunks only once while Multiple
  reads minimize the accesses to the image. Evaluation on a
  3850x3025x3500 brain image shows that our algorithms perform
  similarly to the optimal configuration (slice merging) provided that
  enough memory is available, which solves our initial problem.  We
  conclude that splitting and merging large 3D images can be done
  efficiently based on simple data formats.
\end{abstract}

@@

@@

\begin{figure}
  \includegraphics[width=0.45\columnwidth]{figures/model-big-brain-comparison.pdf}
  \includegraphics[width=0.45\columnwidth]{figures/model-big-brain-comparison-moreblocks.pdf}
  \caption{Number of seeks for Cluster reads vs Multiple reads when
    Cluster reads merge complete block rows (case 2), for D=3458 and
    b=2. Left: n=125; Right: n=64,000.}
  % Gnuplot file: scripts/model/model.gnplt
  \label{fig:model-comparison}
\end{figure}

@@

\subsection{Hardware}

@@

Figure~\ref{fig:slices-vs-blocks} shows the merging time for Slices vs
Blocks, on SSD and HDD. Merging with Blocks takes about twice the time
of Slices, which motivates the need for improved algorithms to merge
blocks. As expected, seek time is mainly responsible for this
difference, although differences in read times are also observed,
perhaps due to a caching effect at the disk or OS level. This behavior is
reproduced in SSD and HDD, which shows that SSDs might not be the only
solution to reduce seek time.  \todo{On figure: check that breakdowns
  add up to total time. Harmonize time formats. Harmonize colors (all
  slices in purple, all blocks in green). If time permits, try
  histogram for breakdown, on same figure than total time.}

@@

\begin{figure}[h]
  \centering
  \includegraphics[width=0.45\columnwidth]{figures/benchmark-buff-slices/buff-slices-breakdown-ssd.pdf}
  \hfill
    \includegraphics[width=0.45\columnwidth]{figures/benchmark--buff-slices/buff-slices-breakdown-hdd.pdf}
  \caption{Comparison between Bufered slices, naive blocks and naive slices. Left: SSD. Right: HDD.}
  % Python file: scripts/experiment/generate_avg_var_data.py
  % Gnuplot file: scripts/experiment/compare_breakdown_errbar.gnuplot
\label{fig:buffered-slices}
\end{figure}

@@

\begin{figure}[h]
  \centering
  \includegraphics[width=0.45\columnwidth]{figures/benchmark-creads/creads-breakdown-ssd.pdf}
  \hfill
    \includegraphics[width=0.45\columnwidth]{figures/benchmark-creads/creads-breakdown-hdd.pdf}
  \caption{Comparison between Cluster reads, naive blocks and naive slices. Left: SSD. Right: HDD.}
  % Python file: scripts/experiment/generate_avg_var_data.py
  % Gnuplot file: scripts/experiment/compare_breakdown_errbar.gnuplot
\label{fig:cluster-reads}
\end{figure}

@@

\begin{figure}[h]
  \centering
  \includegraphics[width=0.45\columnwidth]{figures/benchmark-mreads/mreads-breakdown-ssd.pdf}
  \hfill
    \includegraphics[width=0.45\columnwidth]{figures/benchmark-mreads/mreads-breakdown-hdd.pdf}
  \caption{Comparison between Multiple reads, naive blocks and naive slices. Left: SSD. Right: HDD.}
  % Python file: scripts/experiment/generate_avg_var_data.py
  % Gnuplot file: scripts/experiment/compare_breakdown_errbar.gnuplot
\label{fig:multiple-reads}
\end{figure}

\begin{figure}[h]
  \centering
  \includegraphics[width=0.70\columnwidth]{figures/number-of-seeks.pdf}
  \hfill
  \caption{Number of seeks for all algorithms, comparing modeled and
    experimental values.}
  % Gnuplot file: scripts/number_of_seeks.gnuplot
\label{fig:number-of-seeks}
\end{figure}
